Collocated Deduplication with Fault Isolation for Virtual Machine Snapshot Backup

### 2013-09-03 snapshot_system/fast14_architecture.tex
fix citations

\section{Architecture and Implementation Details}
\label{sect:arch}
Our system runs on a cluster of Linux machines with Xen-based VMs.
A distributed file system (DFS) manages  the physical disk storage and we use 
an open source DFS called QFS~\cite{QFS}. 
All data needed for VM services, such as virtual disk images used by runtime VMs,
and snapshot data for backup purposes, reside in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

@@

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to virtual disk.
Every bit in the bitmap represents a fix-sized (2MB) region called a \textit{segment}, indicating whether the segment
is modified since last backup. 
Segments are further divided into variable-sized chunks (average 4KB) 
using a content-based chunking algorithm~\cite{?}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
[What happens with modification during snapshot backup stage.]

@@

\subsection{A VM-centric snapshot store for backup data}
We build the snapshot storage on the top of a distributed file system.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
There is also a special store containing all PDS chunks shared among different VMs.
As shown in Fig.\ref{fig:as_arch}, we explain the data structure of snapshot stores as follows.

@@

Each container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and store the entire
group as a unit after a compression.
When accessing a particular chunk, its chunk group is retrieved from the storage
and uncompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Sampling,FoundationPaper},
retrieval of  a data chunk  group naturally works well with prefetching. 
A  typical chunk group contains 100 to 1000 chunks, with an average size of 
200-600 chunks.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used through the VM snapshot metadata recipe.
That requires a scanning of all chunks in a VM; however it is a VM-specific procedure and thus
the cost is relatively small compared to the traditional mark-sweep procedure~\cite{?} which scans snapshot 
chunks from all VMs.
For example,
consider each reference pointer consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 90MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

>>>>

\section{Architecture and Implementation Details}
\label{sect:architecture}
Our system runs on a cluster of Linux machines with Xen-based VMs.
A distributed file system (DFS) manages  the physical disk storage and we use 
an open source DFS called QFS~\cite{QFS}. 
All data needed for VM services, such as virtual disk images used by runtime VMs,
and snapshot data for backup purposes, reside in this distributed file system. 
One physical node hosts tens of VMs, each of which accesses its virtual machine disk image through the
virtual block device driver (called TapDisk\cite{Warfield2005} in Xen).

@@

We use the virtual device driver in Xen that employs a bitmap to track the changes 
that have been made to virtual disk.
Every bit in the bitmap represents a fix-sized (2MB) region called a \textit{segment}, indicating whether the segment
is modified since last backup. 
Segments are further divided into variable-sized chunks (average 4KB) 
using a content-based chunking algorithm~\cite{frame05}, 
which brings the opportunity of fine-grained deduplication.
When the VM issues a disk write, the dirty bit for the corresponding segment is set
and this indicates such a segments needs to be checked during snapshot backup. 
After the snapshot backup is finished, the driver resets the dirty bit map to a clean state.
[What happens with modification during snapshot backup stage.]

@@

\subsection{A VM-centric snapshot store for backup data}
\label{sect:store}
We build the snapshot storage on the top of a distributed file system.
Following the VM-centric idea for the purpose of fault isolation,
each VM has its own snapshot store, containing new data chunks which are considered
to be non-duplicates.
There is also a special store containing all PDS chunks shared among different VMs.
As shown in Fig.\ref{fig:as_arch}, we explain the data structure of snapshot stores as follows.

@@

Each container is further divided into a set of chunk data groups. Each chunk group is composed of
a set of data chunks and is the basic unit in data access and retrieval. 
In writing a chunk during backup, the system accumulates data chunks and store the entire
group as a unit after a compression.
When accessing a particular chunk, its chunk group is retrieved from the storage
and uncompressed. Given the high spatial locality and usefulness of prefetching  in 
snapshot chunk accessing~\cite{Guo2011,foundation08},
retrieval of  a data chunk  group naturally works well with prefetching. 
A  typical chunk group contains 100 to 1000 chunks, with an average size of 
200-600 chunks.

@@

\item {\bf Periodic repair of leakage}.
%[exlpain why second Bloom filter, why scan append store]
Leakage repair is conducted periodically to fix the above approximation error.
This procedure compares the live chunks for each VM with what are truly used through the VM snapshot metadata recipe.
That requires a scanning of all chunks in a VM; however it is a VM-specific procedure and thus
the cost is relatively small compared to the traditional mark-sweep procedure~\cite{Guo2011} which scans snapshot 
chunks from all VMs.
For example,
consider each reference pointer consumes 8 bytes plus  1 mark bit, a VM that has 40GB backup data with about
10 million chunks will need less than 90MB of memory to complete a VM-specific mark-sweep process.
\end{itemize}

### 2013-09-03 snapshot_system/fast14_conclusion.tex
fix citations

\section{Conclusion}
\label{sect:final}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency.
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set.
VM-specific file block packing also reduces inter-VM dependence and enhances fault tolerance.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that
our solution can accomplish 92\% of what complete global
deduplication can do while the availability of snapshots increases substantially with this VM-centric
and a small replication overhead for popular inter-VM chunks.

>>>>

\section{Conclusion}
\label{sect:conclusion}
In this paper we propose a collocated backup service built on
the top of a cloud cluster to reduce network traffic and infrastructure.
The key contribution is a VM-centric deduplication scheme to 
maximize fault isolation while delivering competitive deduplication efficiency.
Inner-VM deduplication localizes backup data dependency and exposes more parallelism  
while cross-VM deduplication with a small common data set.
VM-specific file block packing also reduces inter-VM dependence and enhances fault tolerance.
The design places a special consideration for low-resource usages as a collocated cloud service.
Evaluation using  VM backup data shows that
our solution can accomplish 92\% of what complete global
deduplication can do while the availability of snapshots increases substantially with this VM-centric
and a small replication overhead for popular inter-VM chunks.

### 2013-09-03 snapshot_system/fast14_deduplication.tex
figure revisions
fix citations

\section{VM-centric Snapshot Deduplication}
\label{sect:dedupe}

@@

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{??},  
we conduct additional local inner-VM deduplication by comparing
dirty segment's chunk fingerprints within  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their approximated content signature is the same.
This segment signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot recipe.
When we are processing a dirty segment,
we can easily find similar segments from the
parent snapshot recipe and load their segment recipes and also set a limit on the 
the number of  similar segments to be  searched. 
 Then 
given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
with those in similar segments.  
For example, with a 2MB segment, there are about 500 fingerprints to compare. By limiting at most 10 similar
segments to search,
the amount of memory for maintaining those fingerprints in similar segment s is small.

\item \textbf{Cross-VM deduplication with popular chunks and replication support}
This step accomplishes the standard global fingerprint  comparison as conducted
in the previous work~\cite{??}.
One key observation is that the inner deduplication has removed many of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called PDS (common data set). 
The popularity of a chunk is the number  of data chunks  from different VMs
that are duplicates of this chunk after the inner VM deduplication.
This number can be computed periodically on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk signatures in a distributed shared memory.

@@

\item \textbf{VM-centric file system block management.}
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using an append-store  scheme
and discuss this in details in Section~\ref{sect:arch}.
The system allows all machines conduct the backup in parallel in different machines, and each machine
conducts the backup of one VM at a time, and thus only requires a write  buffer for one VM.

@@

\subsection{Impact on deduplication efficiency}
Choosing the value  $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
Our previous study shows that the popularity of data chunks after inner VM deduplication follows 
a Zipf-like distribution\cite{zipf} and its
exponent $\alpha$ is ranged between 0.65  and  0.7.~\cite{ieeecloud}. 
Table~\ref{tab:symbol} lists parameters used in this analysis.

@@

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={PDS Coverage},
            width=\linewidth,
            height=0.6\linewidth,
            cycle multi list={
                mline\nextlist
                [3 of]mmark*\nextlist
            },
            %cycle list name=mcolor,
            xlabel={Total num. chunks stored (in billions)},
            ylabel={PDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.25)},
            anchor=north},
            ]
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A4] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T4] {figures/cds_coverage.txt};
            \legend{Measured ($\sigma=1\%$),Measured ($\sigma=2\%$),Measured ($\sigma=4\%$),Predicted ($\sigma=1\%$),Predicted ($\sigma=2\%$),Predicted ($\sigma=4\%$)};
            \end{axis}
    \end{tikzpicture}
  \caption{Predicted vs. actual PDS coverage as data size increases.}
  \label{fig:cds-coverage}
\end{figure}

@@

The replication degree of the backup storage 
is $r$ for regular file blocks and $r=3$ is a typical setting in the distributed
file system~\cite{Hadoop,GFS}.
In the VC approach, a special replica degree $r_c$ used for PDS blocks where $r_c>r$. 
The storage cost for VO with full deduplication is $c_u *r$ and for VC, it is
$ k*r_c  + (c_u-k)*r$. Thus The storage cost ratio of $VC$ over $VO$ is 
\[
\sigma \frac{r_c}{r} + 1-\sigma
\]
Since $sigma$ is small,  term $\sigma \frac{r_c}{r}$ is small in the above expression.  
Thus the impact on storage increase is very small when we choose a large $\frac{r_c}{r}$ ratio. 
For example, $\frac{r_c}{r}=2$.

>>>>

\section{VM-centric Snapshot Deduplication}
\label{sect:deduplication}

@@

\begin{figure}[htbp]
  \centering
  \epsfig{file=images/mh_local_dedup.png, width=3in}
  \caption{Min-hash based local deduplication}
  \label{fig:local_dedup}
\end{figure}

Since the best deduplication uses a nonuniform chunk size 
in the average of 4K or 8K~\cite{Jin2009},
we conduct additional local inner-VM deduplication by comparing
dirty segment's chunk fingerprints within  {\em similar} segments from its parent snapshot. 
We define two segments are similar if their approximated content signature is the same.
This segment signature value is defined as the minimum value of all its chunk fingerprints 
computed during backup and is recorded in the snapshot recipe.
When we are processing a dirty segment,
we can easily find similar segments from the
parent snapshot recipe and load their segment recipes and also set a limit on the 
the number of  similar segments to be  searched. 
 Then 
given a set of data chunks within a dirty segment,  we compare  these chunk fingerprints
with those in similar segments.  
For example, with a 2MB segment, there are about 500 fingerprints to compare. By limiting at most 10 similar
segments to search,
the amount of memory for maintaining those fingerprints in similar segment s is small.

\item \textbf{Cross-VM deduplication with popular chunks and replication support}
This step accomplishes the standard global fingerprint comparison as conducted
in the previous work~\cite{WeiZhangIEEE}.
One key observation is that the inner deduplication has removed many of the duplicates.
There are fewer deduplication opportunities across VMs while the memory and network
consumption for global comparison is more expensive.
Thus our approximation is that the global fingerprint comparison only searches for the top $k$
most popular items. This dataset is called PDS (common data set). 
The popularity of a chunk is the number  of data chunks  from different VMs
that are duplicates of this chunk after the inner VM deduplication.
This number can be computed periodically on a weekly basis.
Once the popularity of all data chunks is collected, the system only maintains the top $k$
most popular chunk signatures in a distributed shared memory.

@@

\item \textbf{VM-centric file system block management.}
When a chunk is not detected as a duplicate to any existing chunk, this chunk will be written
to the file system. Since the backend file system typically uses a large block size such as 64MB, each VM will 
accumulate small local chunks. We manage this accumulation process using an append-store  scheme
and discuss this in details in Section~\ref{sect:architecture}.
The system allows all machines conduct the backup in parallel in different machines, and each machine
conducts the backup of one VM at a time, and thus only requires a write  buffer for one VM.

@@

\subsection{Impact on deduplication efficiency}
Choosing the value  $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
Our previous study shows that the popularity of data chunks after inner VM deduplication follows 
a Zipf-like distribution\cite{Breslau1999a} and its
exponent $\alpha$ is ranged between 0.65  and  0.7.~\cite{WeiZhangIEEE}. 
Table~\ref{tab:symbol} lists parameters used in this analysis.

@@

\begin{figure}[htbp]
  \centering
    \begin{tikzpicture}
            \begin{axis}[
            %title={PDS Coverage},
            width=\linewidth,
            height=0.6\linewidth,
            cycle multi list={
                mline\nextlist
                [3 of]mmark*\nextlist
            },
            %cycle list name=mcolor,
            xlabel={Total num. chunks stored (in billions)},
            ylabel={PDS Coverage (\%)},
            %extra y ticks={4.5,5.5,6.5} %to add extra ticks
            mark options=solid,
            %legend pos=outer north east,
            legend columns=2,
            legend style={
                at={(0.5,-0.30)},
            anchor=north},
            ]
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=A4] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T1] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T2] {figures/cds_coverage.txt};
            \addplot table[x expr=\thisrow{InputChunks}/1000000000,y=T4] {figures/cds_coverage.txt};
            \legend{Measured ($\sigma=1\%$),Measured ($\sigma=2\%$),Measured ($\sigma=4\%$),Predicted ($\sigma=1\%$),Predicted ($\sigma=2\%$),Predicted ($\sigma=4\%$)};
            \end{axis}
    \end{tikzpicture}
  \caption{Predicted vs. actual PDS coverage as data size increases.}
  \label{fig:cds-coverage}
\end{figure}

@@

The replication degree of the backup storage 
is $r$ for regular file blocks and $r=3$ is a typical setting in the distributed
file system~\cite{googlefs03,hdfs10}.
In the VC approach, a special replica degree $r_c$ used for PDS blocks where $r_c>r$. 
The storage cost for VO with full deduplication is $c_u *r$ and for VC, it is
$ k*r_c  + (c_u-k)*r$. Thus The storage cost ratio of $VC$ over $VO$ is 
\[
\sigma \frac{r_c}{r} + 1-\sigma
\]
Since $sigma$ is small,  term $\sigma \frac{r_c}{r}$ is small in the above expression.  
Thus the impact on storage increase is very small when we choose a large $\frac{r_c}{r}$ ratio. 
For example, $\frac{r_c}{r}=2$.

### 2013-09-03 snapshot_system/fast14_evaluation.tex
fix citations

\section{Evaluation}
\label{sect:exper}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 5.
Our evaluation objective is to
study the benefit of fault tolerance and   deduplication efficiency  the VC approach and compare it
with an alternative VO design, and assess backup throughput and  resource usage for a large number of VMs.

We will compare our VC approach with
a VO approach  using stateless routing with binning (call it SRB) 
based on~\cite{ParallelDataDomain,Extremebining}
SRB executes a distributed deduplication by routing a data chunk to a machine~\cite{ParallelDataDomain}
using  a min-hash function discussed in \cite{Extremebining}. Once a data chunk is routed to
a machine, this chunk is compared with the fingerprint index of this machine locally.

\subsection{Settings}
We have performed a trace-driven study using a production dataset~\cite{WeiZhangIEEE} from 
Alibaba Aliyun's cloud platform with about 100 machines. 
The VMs  of the sampled  data set   use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, and win2008. 
Each machine hosts upto 25 VMs and each VM keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.
The backup of VM snapshots is completed within a few  hours every night.
Based on our study of production  data,  each VM has about  40GB of storage  data  on average
including OS and user data disk.
All data are divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content chunks ~\cite{similar94,rabin81} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
Popularity of data blocks are collected through global counting and 
and top popular ones ($\delta=1-2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:crossVM}.

@@

\subsection{Deduplication Efficiency}
Figure~\ref{fig:exbin-efficiency-graph} shows the deduplication efficiency for SRB and VC.
Deduplication efficiency is defined as the percent of duplicate chunks
which are detected and deduplicated.
The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
is defined in Table~\ref{tab:symbol}. 
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB
and  the deduplication efficiency can reach over 90\%.
When $\sigma=4\%$, the deduplication efficiency can reach 92\%. 
The loss of efficiency in VC is caused by the restriction of the total physical memory available
in the cluster to support PDS index.  
The loss of efficiency in SRB is caused by the routing of data chunks which restricts the search scope
for global comparison.
In all cases, VC provides similar or better deduplication efficiency than SRB.

>>>>

\section{Evaluation}
\label{sect:evaluation}
We have implemented and evaluated a prototype of our VC scheme on a Linux cluster of machines with
8-core 3.1Ghz AMD FX-8120 and 16 GB RAM. 
Our implementation is based on Alibaba cloud platform~\cite{Aliyun,WeiZhangIEEE}
and the underlying DFS uses  QFS with default replication degree 3 while the PDS replication degree is 5.
Our evaluation objective is to
study the benefit of fault tolerance and   deduplication efficiency  the VC approach and compare it
with an alternative VO design, and assess backup throughput and  resource usage for a large number of VMs.

We will compare our VC approach with
a VO approach  using stateless routing with binning (call it SRB) 
based on~\cite{Dong2011,extreme_binning09}
SRB executes a distributed deduplication by routing a data chunk to a machine~\cite{Dong2011}
using  a min-hash function discussed in \cite{extreme_binning09}. Once a data chunk is routed to
a machine, this chunk is compared with the fingerprint index of this machine locally.

\subsection{Settings}
We have performed a trace-driven study using a production dataset~\cite{WeiZhangIEEE} from 
Alibaba Aliyun's cloud platform with about 100 machines. 
The VMs  of the sampled  data set   use popular operating systems such as 
Debian, Ubuntu, Redhat, CentOS, and win2008. 
Each machine hosts upto 25 VMs and each VM keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.
The backup of VM snapshots is completed within a few  hours every night.
Based on our study of production  data,  each VM has about  40GB of storage  data  on average
including OS and user data disk.
All data are divided into 2 MB fix-sized segments and each segment is divided into 
variable-sized content chunks ~\cite{similar94,rabin81} with an average size of 4KB.
The signature for variable-sized blocks is computed using their SHA-1 hash. 
Popularity of data blocks are collected through global counting and 
and top popular ones ($\delta=1-2\%$)  are  kept in the PDS, as discussed in Section~\ref{sect:store}.

@@

\subsection{Deduplication Efficiency}
Figure~\ref{fig:exbin-efficiency-graph2} shows the deduplication efficiency for SRB and VC.
Deduplication efficiency is defined as the percent of duplicate chunks
which are detected and deduplicated.
The figure also compares several PDS sizes chosen for VC. ``$\sigma=2\%$'' 
is defined in Table~\ref{tab:symbol}. 
%we allocate the space of distributed shared memory  to accommodate 2\%
%of data chunks shared among VMs, as defined in 
With $\sigma=2\%$, memory usage for PDS index lookup per machine is about 100MB
and  the deduplication efficiency can reach over 90\%.
When $\sigma=4\%$, the deduplication efficiency can reach 92\%. 
The loss of efficiency in VC is caused by the restriction of the total physical memory available
in the cluster to support PDS index.  
The loss of efficiency in SRB is caused by the routing of data chunks which restricts the search scope
for global comparison.
In all cases, VC provides similar or better deduplication efficiency than SRB.
